### 2024-09-24 paper/paper.qmd
first draft, show the basic trend of the data, analysis in word needed, futher clean/test/simuate needed

---
title: "Deaths of People Experiencing Homelessness In Toronto"
subtitle: "My subtitle if needed"
author: 
  - Sakura Hu
thanks: "Code and data are available at: https://open.toronto.ca/dataset/deaths-of-people-experiencing-homelessness/."
date: "2024-09-23"
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-yearly-deaths
#| fig-cap: Yearly Distribution of Deaths of People Experiencing Homelessness
#| echo: false

analyze_data <- read_csv("../data/raw_data/raw_data.csv")

```

```{r}
# Reshape the data to long format and count occurrences for each bias
bias_counts <- cleaned_data %>%
  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
  mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
  group_by(Bias_Type) %>%
  summarise(Count = sum(Value))

# Create a bar chart
ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Bias Type", y = "Count of Occurrences", title = "Count of Occurrences Based on Different Bias Types") +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

# Filter out "None" from the RACE_BIAS column and count occurrences
race_bias_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  count(RACE_BIAS)

# Create the bar chart
ggplot(race_bias_counts, aes(x = reorder(RACE_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Race Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Count occurrences by OCCURRENCE_YEAR without filtering
year_counts <- cleaned_data %>%
  count(OCCURRENCE_YEAR)

# Create the bar chart
ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Year") +
  scale_x_discrete(drop = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Filter out "None" and create a new column for Race Bias categorization
race_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  mutate(RACE_CATEGORY = ifelse(RACE_BIAS == "Black", "Black", "Other"))

# Count occurrences by OCCURRENCE_YEAR and RACE_CATEGORY
year_race_counts <- race_counts %>%
  count(OCCURRENCE_YEAR, RACE_CATEGORY)

# Create the stacked bar chart
ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n, fill = RACE_CATEGORY)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
  scale_fill_manual(values = c("Black" = "steelblue", "Other" = "orange")) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Year and Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the ETHNICITY_BIAS column and count occurrences
ethnicity_bias_counts <- cleaned_data %>%
  filter(ETHNICITY_BIAS != "None") %>%
  count(ETHNICITY_BIAS)

# Create the bar chart
ggplot(ethnicity_bias_counts, aes(x = reorder(ETHNICITY_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "ethnicity Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by ethnicity Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the LANGUAGE_BIAS column and count occurrences
language_bias_counts <- cleaned_data %>%
  filter(LANGUAGE_BIAS != "None") %>%
  count(LANGUAGE_BIAS)

# Create the bar chart
ggplot(language_bias_counts, aes(x = reorder(LANGUAGE_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Language Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Language Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Filter out "None" from the RELIGION_BIAS column and count occurrences
religion_bias_counts <- cleaned_data %>%
  filter(RELIGION_BIAS != "None") %>%
  count(RELIGION_BIAS)

# Create the bar chart
ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Religion Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Language Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Filter out "None" from the SEXUAL_ORIENTATION_BIAS column and count occurrences
sexual_ori_bias_counts <- cleaned_data %>%
  filter(SEXUAL_ORIENTATION_BIAS != "None") %>%
  count(SEXUAL_ORIENTATION_BIAS)

# Create the bar chart
ggplot(sexual_ori_bias_counts, aes(x = reorder(SEXUAL_ORIENTATION_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Sexual Orientation Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Sexual Orientation Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the GENDER_BIAScolumn and count occurrences
gender_bias_counts <- cleaned_data %>%
  filter(GENDER_BIAS != "None") %>%
  count(GENDER_BIAS)

# Create the bar chart
ggplot(gender_bias_counts, aes(x = reorder(GENDER_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Gender Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Gender Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-09-24 scripts/00-simulate_data.R

@@ -1,19 +1,35 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates data for Hate Crimes analysis
+# Author: Sakura Hu
+# Date: 20 September 2024
+# Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(1009201917)
+
+# Define the number of records to simulate
+number_of_records <- 84
+
+# Define the simulation parameters
+years <- seq(2017, 2023)  # Years for the simulation
+months <- month.name  # Month names from January to December
+
+# Create the simulated data
+simulated_data <- tibble(
+  id = 1:number_of_records,  # Unique row identifier
+  Year_of_death = sample(years, number_of_records, replace = TRUE),  # Randomly sample years
+  Month_of_death = sample(months, number_of_records, replace = TRUE),  # Randomly sample months
+  Count = rpois(number_of_records, lambda = 13)  # Randomly generate counts using Poisson distribution
+)
 
 
-
+#### Write_csv
+write_csv(simulated_data, file = "data/raw_data/simulated.csv")

### 2024-09-24 scripts/01-download_data.R

@@ -1,26 +1,47 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Sakura Hu
+# Date: 20 September 2024
+# Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
+# Install opendatatoronto if not already installed
+install.packages("opendatatoronto")
+
+# Load the required library
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
 
+# get package
+package <- show_package("hate-crimes-open-data")
+package
+
+# get all resources for this package
+resources <- list_package_resources("hate-crimes-open-data")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data
+#### Save data ####
+write_csv(data, "data/raw_data/raw_data.csv") 
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change the_raw_data to whatever name you assigned when you downloaded it.
+write_csv(data, "data/raw_data.csv") 
 #### Download data ####
 # [...ADD CODE HERE TO DOWNLOAD...]
-
 
 
 #### Save data ####
 # [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         
+write_csv(raw_data, "inputs/data/raw_data.csv") 

### 2024-09-24 scripts/02-data_cleaning.R

@@ -7,38 +7,28 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
+#### Preamble ####
+# Purpose: Cleans the raw hate crime data recorded by observers
+# Author: Rohan Alexander
+# Date: 6 April 2023
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites: R packages tidyverse, dplyr must be installed
+# Any other information needed? This script processes the hate crime dataset.
+
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+# Replace "NO" with "None" in the specified columns
+cleaned_data <- raw_data %>%
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+    AGE_BIAS = ifelse(AGE_BIAS == "NO", "None", AGE_BIAS),
+    MENTAL_OR_PHYSICAL_DISABILITY = ifelse(MENTAL_OR_PHYSICAL_DISABILITY == "NO", "None", MENTAL_OR_PHYSICAL_DISABILITY)
+  )
+
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-24 scripts/03-test_data.R

@@ -1,15 +1,26 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Sanity check of the data
+# Author: Sakura Hu
+# Date: 20 September 2024
+# Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Need to have simulated data
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+simulated_data <- read_csv("data/raw_data/simulated.csv")
+
+# Test for negative numbers
+simulated_data$Count |> min() <= 0
+
+# Test for NAs
+all(is.na(simulated_data$Count))
+
+# Test for unqiue id
+
+# Test for valid

### 2024-09-24 paper/paper.qmd
first draft for abstract and intro

@@
 ---
-title: "Deaths of People Experiencing Homelessness In Toronto"
-subtitle: "My subtitle if needed"
+title: "Analyzing Patterns of Hate Crimes: Insights from Toronto's Hate Crime Dataset"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper explores the issue of hate crimes in Toronto and their underlying causes, highlighting the urgent need for societal awareness and intervention.
+Utilizing a dataset of hate crime incidents, we employed statistical methods to analyze the counts attributed to various bias categories.
+Our findings reveal that the majority of hate crimes are directed toward Jewish, female, and Black individuals, with a concerning trend of increasing occurrences over the years.
+These results underscore the importance of understanding hate crimes to inform policies and initiatives aimed at fostering inclusivity and reducing discrimination in our communities."
 format: pdf
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Hate crimes are criminal offenses directed at individuals or property, driven by bias, prejudice, or hatred based on factors such as race, ethnicity, language, religion, gender, sexual orientation, disability, or other similar characteristics.
+These crimes are a serious societal issue, as they target specific groups and reinforce divisions.
+In recent years, cities like Toronto have witnessed a rise in reported hate crimes, prompting a need to analyze the underlying causes and trends.
+Understanding the motivations behind hate crimes and identifying the most affected communities is critical for developing targeted prevention strategies and fostering a more inclusive society.
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper utilizes data on hate crimes reported in Toronto to explore the causes behind these offenses.
+Specifically, it investigates the most frequent motivations for hate crimes, including race, ethnicity, religion, sexual orientation, and gender identity.
+By analyzing which demographic groups are most often victimized, this paper aims to shed light on patterns of bias in hate crimes and how these trends have evolved over time.
+The analysis reveals that Black, Jewish, and female individuals are disproportionately targeted, and the occurrence of hate crimes has been increasing each year.
+
+The importance of this study lies in its ability to inform policymakers, law enforcement, and the public about the prevalence and nature of hate crimes in Toronto.
+Addressing the root causes of these offenses is essential for reducing their occurrence and protecting vulnerable communities.
+
+The remainder of this paper is structured as follows: Section 2 discusses the data sources and variables, as well as the process of cleaning and preparing the data for analysis.
+Section 3 presents the results, including visualizations of key trends, while Section 4 provides a discussion of the findings.
+Finally, Section 5 includes appendices with additional details and supporting material.
 

### 2024-09-26 paper/paper.qmd
second draft, analysis + final check + sketch + reference needed

@@
 ---
-title: "Analyzing Patterns of Hate Crimes: Insights from Toronto's Hate Crime Dataset"
+title: "Analyzing Patterns of Hate Crimes"
+subtitle: "WHAT DID YOU FIND?"
 author: 
   - Sakura Hu
-thanks: "Code and data are available at: https://open.toronto.ca/dataset/deaths-of-people-experiencing-homelessness/."
+thanks: "https://open.toronto.ca/dataset/hate-crimes-open-data/"
 date: "2024-09-23"
@@
 library(dplyr)
+library(knitr)
 ```
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
 
-```{r}
-#| label: fig-yearly-deaths
-#| fig-cap: Yearly Distribution of Deaths of People Experiencing Homelessness
-#| echo: false
+cleaned_data <- read_csv("../data/analysis_data/analysis_data.csv")
+# Summarize the counts of occurrences for different bias categories by year
+bias_summary_by_year <- cleaned_data %>%
+  group_by(OCCURRENCE_YEAR) %>%
+  summarize(
+    race_bias_count = sum(RACE_BIAS != "None"),
+    ethnicity_bias_count = sum(ETHNICITY_BIAS != "None"),
+    religion_bias_count = sum(RELIGION_BIAS != "None"),
+    gender_bias_count = sum(GENDER_BIAS != "None"),
+    sexual_orientation_bias_count = sum(SEXUAL_ORIENTATION_BIAS != "None"),
+    disability_bias_count = sum(MENTAL_OR_PHYSICAL_DISABILITY != "None")
+  )
 
-analyze_data <- read_csv("../data/raw_data/raw_data.csv")
-
+# Display the summary
+bias_summary_by_year %>% 
+  kable(
+    col.names = c("Year", "Race", "Ethnicity", "Religion", "Gender", "Sexual_Orientation", "Disability"),
+    # digits = 1
+  )
 ```
@@
 ```{r}
+#| label: fig-occurence-by-bias
+#| fig-cap: Count of Occurrences Based on Different Bias Types
+#| echo: false
 # Reshape the data to long format and count occurrences for each bias
@@
   theme_minimal() +
-  labs(x = "Bias Type", y = "Count of Occurrences", title = "Count of Occurrences Based on Different Bias Types") +
+  labs(x = "Bias Type", y = "Count of Occurrences") +
   scale_fill_brewer(palette = "Set3") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+#| label: fig-yearly-occurrence
+#| fig-cap: Occurrences of Hate Crimes by Year
+#| echo: false
+# Count occurrences by OCCURRENCE_YEAR without filtering
+year_counts <- cleaned_data %>%
+  count(OCCURRENCE_YEAR)
+
+# Create the bar chart
+ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
+  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Year of Occurrence", y = "Number of Occurrences") +
+  scale_x_discrete(drop = FALSE) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 ```{r}
+#| label: fig-religion-bias
+#| fig-cap: Occurrences of Hate Crimes by Religion Bias
+#| echo: false
+# Filter out "None" from the RELIGION_BIAS column and count occurrences
+religion_bias_counts <- cleaned_data %>%
+  filter(RELIGION_BIAS != "None") %>%
+  count(RELIGION_BIAS)
 
+# Create the bar chart
+ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
+  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Religion Bias Categories", y = "Number of Occurrences") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r}
+#| label: fig-Jewish-bias
+#| fig-cap: Occurrences of Hate Crimes by Year and Race Bias
+#| echo: false
+# Filter out "None" and create a new column for Race Bias categorization
+race_counts <- cleaned_data %>%
+  filter(RELIGION_BIAS != "None") %>%
+  mutate(RELIGION_CATEGORY = ifelse(RELIGION_BIAS == "Jewish", "Jewish", "Other"))
+
+# Count occurrences by OCCURRENCE_YEAR and RACE_CATEGORY
+year_religion_counts <- race_counts %>%
+  count(OCCURRENCE_YEAR, RELIGION_CATEGORY)
+
+# Calculate percentage for each category within each year
+year_religion_counts <- year_religion_counts %>%
+  group_by(OCCURRENCE_YEAR) %>%
+  mutate(percentage = n / sum(n) * 100)
+
+# Create the stacked bar chart
+ggplot(year_religion_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RELIGION_CATEGORY)) +
+  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
+  scale_fill_manual(values = c("Jewish" = "steelblue", "Other" = "orange")) +
+  theme_minimal() +
+  labs(x = "Year of Occurrence", y = "Number of Occurrences") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_text(aes(label = scales::percent(percentage / 100)), 
+            position = position_stack(vjust = 0.5), 
+            color = "white", size = 4)  # Adjust size and color as needed
+```
+
+```{r}
+#| echo: false
 # Filter out "None" from the RACE_BIAS column and count occurrences
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-```
 ```{r}
-# Count occurrences by OCCURRENCE_YEAR without filtering
-year_counts <- cleaned_data %>%
-  count(OCCURRENCE_YEAR)
-
-# Create the bar chart
-ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Year") +
-  scale_x_discrete(drop = FALSE) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
+#| echo: false
 # Filter out "None" and create a new column for Race Bias categorization
@@
 
-# Create the stacked bar chart
-ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n, fill = RACE_CATEGORY)) +
+# Calculate percentage for each category within each year
+year_race_counts <- year_race_counts %>%
+  group_by(OCCURRENCE_YEAR) %>%
+  mutate(percentage = n / sum(n) * 100)
+
+# Create the stacked bar chart with percentage labels
+ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RACE_CATEGORY)) +
   geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
@@
   theme_minimal() +
-  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Year and Race Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-```{r}
-# Filter out "None" from the ETHNICITY_BIAS column and count occurrences
-ethnicity_bias_counts <- cleaned_data %>%
-  filter(ETHNICITY_BIAS != "None") %>%
-  count(ETHNICITY_BIAS)
-
-# Create the bar chart
-ggplot(ethnicity_bias_counts, aes(x = reorder(ETHNICITY_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "ethnicity Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by ethnicity Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-```{r}
-# Filter out "None" from the LANGUAGE_BIAS column and count occurrences
-language_bias_counts <- cleaned_data %>%
-  filter(LANGUAGE_BIAS != "None") %>%
-  count(LANGUAGE_BIAS)
-
-# Create the bar chart
-ggplot(language_bias_counts, aes(x = reorder(LANGUAGE_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Language Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Language Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
-# Filter out "None" from the RELIGION_BIAS column and count occurrences
-religion_bias_counts <- cleaned_data %>%
-  filter(RELIGION_BIAS != "None") %>%
-  count(RELIGION_BIAS)
-
-# Create the bar chart
-ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Religion Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Language Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-```{r}
-# Filter out "None" from the SEXUAL_ORIENTATION_BIAS column and count occurrences
-sexual_ori_bias_counts <- cleaned_data %>%
-  filter(SEXUAL_ORIENTATION_BIAS != "None") %>%
-  count(SEXUAL_ORIENTATION_BIAS)
-
-# Create the bar chart
-ggplot(sexual_ori_bias_counts, aes(x = reorder(SEXUAL_ORIENTATION_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Sexual Orientation Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Sexual Orientation Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-```{r}
-# Filter out "None" from the GENDER_BIAScolumn and count occurrences
-gender_bias_counts <- cleaned_data %>%
-  filter(GENDER_BIAS != "None") %>%
-  count(GENDER_BIAS)
-
-# Create the bar chart
-ggplot(gender_bias_counts, aes(x = reorder(GENDER_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Gender Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Gender Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  labs(x = "Year of Occurrence", y = "Percentage of Occurrences", 
+       title = "Percentage of Hate Crimes by Year and Race Bias") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_text(aes(label = scales::percent(percentage / 100)), 
+            position = position_stack(vjust = 0.5), 
+            color = "white", size = 4)  # Adjust size and color as needed
 ```
@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
 Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 
@@
 Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-09-26 scripts/00-simulate_data.R

@@ -16,19 +16,23 @@ library(tidyverse)
 set.seed(1009201917)
 
 # Define the number of records to simulate
-number_of_records <- 84
+n <- 100
 
-# Define the simulation parameters
-years <- seq(2017, 2023)  # Years for the simulation
-months <- month.name  # Month names from January to December
+# Simulating OCCURRENCE_YEAR (e.g., from 2018 to 2023)
+occurrence_year <- sample(2018:2023, n, replace = TRUE)
 
-# Create the simulated data
-simulated_data <- tibble(
-  id = 1:number_of_records,  # Unique row identifier
-  Year_of_death = sample(years, number_of_records, replace = TRUE),  # Randomly sample years
-  Month_of_death = sample(months, number_of_records, replace = TRUE),  # Randomly sample months
-  Count = rpois(number_of_records, lambda = 13)  # Randomly generate counts using Poisson distribution
-)
+simulated_data <- 
+  tibble(
+    OCCURRENCE_YEAR = occurrence_year,
+    RACE_BIAS = sample(unique(cleaned_data$RACE_BIAS), n, replace = TRUE),
+    ETHNICITY_BIAS = sample(unique(cleaned_data$ETHNICITY_BIAS), n, replace = TRUE),
+    RELIGION_BIAS = sample(unique(cleaned_data$RELIGION_BIAS), n, replace = TRUE),
+    SEXUAL_ORIENTATION_BIAS = sample(unique(cleaned_data$SEXUAL_ORIENTATION_BIAS), n, replace = TRUE),
+    GENDER_BIAS = sample(unique(cleaned_data$GENDER_BIAS), n, replace = TRUE)
+  )
+
+# Check the first few rows of the simulated data
+head(simulated_data)
 
 
 #### Write_csv

### 2024-09-26 scripts/02-data_cleaning.R

@@ -1,17 +1,8 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Preamble ####
 # Purpose: Cleans the raw hate crime data recorded by observers
-# Author: Rohan Alexander
-# Date: 6 April 2023
-# Contact: rohan.alexander@utoronto.ca
+# Author: Sakura Hu
+# Date: 23 September 2024
+# Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: R packages tidyverse, dplyr must be installed
 # Any other information needed? This script processes the hate crime dataset.
@@ -27,8 +18,10 @@ cleaned_data <- raw_data %>%
   mutate(
     AGE_BIAS = ifelse(AGE_BIAS == "NO", "None", AGE_BIAS),
     MENTAL_OR_PHYSICAL_DISABILITY = ifelse(MENTAL_OR_PHYSICAL_DISABILITY == "NO", "None", MENTAL_OR_PHYSICAL_DISABILITY)
-  )
-
+  ) %>%
+  # Remove specified columns
+  select(-NEIGHBOURHOOD_140, -HOOD_140, -NEIGHBOURHOOD_158, -HOOD_158, -DIVISION, -REPORTED_TIME, 
+         -OCCURRENCE_TIME, -REPORTED_YEAR, -REPORTED_DATE, -AGE_BIAS, -MULTIPLE_BIAS, -LOCATION_TYPE)
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-27 paper/paper.qmd
third draft, still need: futher analysis, reference, final check

@@
 ---
-title: "Analyzing Patterns of Hate Crimes"
-subtitle: "WHAT DID YOU FIND?"
+title: "Analyzing Patterns of Hate Crimes in Toronto"
+subtitle: "majority of hate crimes are directed toward Jewish "
 author: 
@@
 thanks: "https://open.toronto.ca/dataset/hate-crimes-open-data/"
-date: "2024-09-23"
+date: "2024-09-27"
 date-format: long
@@
 Utilizing a dataset of hate crime incidents, we employed statistical methods to analyze the counts attributed to various bias categories.
-Our findings reveal that the majority of hate crimes are directed toward Jewish, female, and Black individuals, with a concerning trend of increasing occurrences over the years.
+Our findings reveal that the majority of hate crimes are directed toward Jewish individuals, with a concerning trend of increasing occurrences over the years.
 These results underscore the importance of understanding hate crimes to inform policies and initiatives aimed at fostering inclusivity and reducing discrimination in our communities."
@@
 # Data {#sec-data}
+## Data Overview
+
+The dataset for this analysis is sourced from Open Data Toronto, specifically from the Toronto Police Service’s verified Hate Crime occurrences.
+It records all confirmed hate crimes that have been investigated by the Hate Crime Unit since 2018, categorized by their reported date.
+Notably, the dataset only includes verified hate crimes and excludes incidents that were unfounded or classified merely as hate incidents.
+
+The dataset serves an important societal purpose, providing communities with detailed information to enhance public safety and awareness about hate crimes in Toronto.
+The data categories, referred to as bias categories, capture various dimensions of prejudice that are present in hate crimes, such as age, mental or physical disability, race, ethnicity, language, religion, sexual orientation, and gender.
+These categories allow researchers to analyze the prevalence of bias in different forms, offering a structured view of hate crime patterns over time.
+
+This dataset is invaluable for understanding hate crimes within the city, but it also comes with ethical considerations.
+Since each data point represents a real-world event, we must recognize the harm experienced by individuals and communities due to these crimes.
+Additionally, it’s important to acknowledge that the dataset focuses on confirmed hate crimes, which means it only represents cases where sufficient evidence was gathered.
+While this strengthens the reliability of the data, it also limits the scope of the analysis to known and proven crimes.
+Furthermore, the geographic locations of the incidents are aggregated to the neighborhood level to protect the privacy of the victims and communities involved.
+
+The data is composed of several key variables relevant to this analysis:
+
+- **EVENT_UNIQUE_ID**: A unique identifier for each offense.
+- **OCCURRENCE_YEAR**: The year in which the offense occurred.
+- **MENTAL_OR_PHYSICAL_DISABILITY**: Whether the offense was based on the suspect’s perception of the victim's mental or physical disability.
+- **RACE_BIAS**: The race of the victim as perceived by the suspect, which was a basis for the hate crime.
+- **ETHNICITY_BIAS**: The ethnicity of the victim as perceived by the suspect, contributing to the hate crime.
+- **LANGUAGE_BIAS**: The language spoken by the victim as perceived by the suspect, used as a bias basis for the crime.
+- **RELIGION_BIAS**: The victim’s religion as perceived by the suspect, which motivated the crime.
+- **SEXUAL_ORIENTATION_BIAS**: The sexual orientation of the victim as perceived by the suspect, serving as a basis for the hate crime.
+- **GENDER_BIAS**: The gender of the victim as perceived by the suspect.
+- **LOCATION_TYPE**: Location type of the offence.
+
+Alternative datasets were considered but were not used in this analysis.
+This particular dataset was chosen because it is published by the Toronto Police Service and provides verified, official data on hate crimes, making it more reliable than other possible datasets from unofficial sources.
+Including additional datasets could introduce inconsistency or unreliability, so this analysis focuses solely on this verified source.
+
+## Data Cleaning
+The dataset underwent several cleaning steps to ensure its relevance and consistency for analysis.
+Several variables were removed as they were either irrelevant to the current analysis or did not add value.
+These included neighborhood-level geographic identifiers (e.g., NEIGHBOURHOOD_140, NEIGHBOURHOOD_158), reporting times and dates (REPORTED_TIME, REPORTED_DATE), and any variables related to the classification or reporting of multiple biases.
+Specifically, AGE_BIAS was removed because it contained only the constant value “None,” suggesting that hate crimes in this dataset are rarely motivated solely by the victim’s age.
+
+Additionally, the MENTAL_OR_PHYSICAL_DISABILITY variable, which initially used "No" to denote a lack of bias, was modified to use "None" instead to maintain consistency with the other bias columns.
+No new variables were constructed, but this cleaning process ensured that the dataset was aligned and ready for analysis focused on the key bias categories.
+
+To handle the NAs in the location type column, rows with NAs location type is removed when doing relative analysis, it is not removed when analysis is irrelavant to location type.
+
+## Data Analysis
 
@@
 #| message: false
+#| label: tbl-summary
+#| tbl-cap: "Summary Statistics"
 
 cleaned_data <- read_csv("../data/analysis_data/analysis_data.csv")
+
 # Summarize the counts of occurrences for different bias categories by year
@@
     disability_bias_count = sum(MENTAL_OR_PHYSICAL_DISABILITY != "None")
+  ) %>%
+  mutate(total_by_year = rowSums(across(starts_with("race_bias_count"):starts_with("disability_bias_count"))))
+
+# Convert OCCURRENCE_YEAR to character
+bias_summary_by_year <- bias_summary_by_year %>%
+  mutate(OCCURRENCE_YEAR = as.character(OCCURRENCE_YEAR))
+
+# Add a row to calculate the total for each bias category across all years
+total_by_category <- bias_summary_by_year %>%
+  summarize(
+    OCCURRENCE_YEAR = "Total",
+    race_bias_count = sum(race_bias_count),
+    ethnicity_bias_count = sum(ethnicity_bias_count),
+    religion_bias_count = sum(religion_bias_count),
+    gender_bias_count = sum(gender_bias_count),
+    sexual_orientation_bias_count = sum(sexual_orientation_bias_count),
+    disability_bias_count = sum(disability_bias_count),
+    total_by_year = sum(total_by_year)
   )
 
-# Display the summary
-bias_summary_by_year %>% 
+# Combine the summary by year with the total row
+bias_summary_with_total <- bind_rows(bias_summary_by_year, total_by_category)
+
+# Display the summary table with totals
+bias_summary_with_total %>%
   kable(
-    col.names = c("Year", "Race", "Ethnicity", "Religion", "Gender", "Sexual_Orientation", "Disability"),
-    # digits = 1
+    col.names = c("Year", "Race", "Ethnicity", "Religion", "Gender", "Sexual Orientation", "Disability", "Total"),
+    format = "pipe" # Use pipe format for compatibility
   )
+
 ```
+@tbl-summary provides a detailed summary statistic of the cleaned data, illustrating the number of hate crimes committed each year, categorized by specific bias types, along with the total occurrences for each category and the overall total of hate crimes annually.
+The table indicates that, on average, there are approximately 250 hate crime incidents reported each year in Toronto over the past six years.
+Notably, incidents motivated by religious bias consistently represent the highest proportion among the various bias types, underscoring a significant area of concern for community safety and policy interventions.
+This data highlights not only the prevalence of hate crimes but also the specific biases that warrant focused attention and resources for prevention and support initiatives.
 
 ```{r}
-#| label: fig-occurence-by-bias
-#| fig-cap: Count of Occurrences Based on Different Bias Types
+#| label: fig-yearly-occurrence
+#| fig-cap: "Occurrences of Hate Crimes by Year"
 #| echo: false
-# Reshape the data to long format and count occurrences for each bias
-bias_counts <- cleaned_data %>%
-  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
-  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
-  mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
-  group_by(Bias_Type) %>%
-  summarise(Count = sum(Value))
 
-# Create a bar chart
-ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
-  geom_bar(stat = "identity") +
-  theme_minimal() +
-  labs(x = "Bias Type", y = "Count of Occurrences") +
-  scale_fill_brewer(palette = "Set3") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
-#| label: fig-yearly-occurrence
-#| fig-cap: Occurrences of Hate Crimes by Year
-#| echo: false
 # Count occurrences by OCCURRENCE_YEAR without filtering
@@
 
+@fig-yearly-occurrence illustrates the annual occurrences of hate crimes from 2018 to 2023.
+The data indicates a clear upward trend in the number of reported hate crimes over this period.
+Starting from approximately 154 occurrences in 2018, there is a steady increase each year, culminating in a significant rise to over 392 occurrences in 2023.
+
+This trend may suggest a growing prevalence of hate crimes within the community, possibly influenced by various social, political, and economic factors.
+Additionally, it could reflect an increase in public awareness and reporting of such incidents, as societal attitudes toward hate crimes continue to evolve.
+This increase raises important questions about the underlying causes and necessitates further investigation into the factors contributing to the rise in these offenses.
+
+Overall, the increasing trend in hate crime occurrences highlights the importance of ongoing efforts in awareness, prevention, and intervention strategies to address the root causes of hate-driven behavior in society.
+
+```{r}
+#| label: fig-occurrence-by-bias
+#| fig-cap: "Count of Occurrences Based on Different Bias Types"
+#| echo: false
+#| fig.width: 8
+#| fig.height: 5
+
+# Reshape the data to long format and count occurrences for each bias
+bias_counts <- cleaned_data %>%
+  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
+  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
+  mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
+  group_by(Bias_Type) %>%
+  summarise(Count = sum(Value))
+
+# Modify Bias_Type for better readability
+bias_counts <- bias_counts %>%
+  mutate(Bias_Type = case_when(
+    Bias_Type == "RACE_BIAS" ~ "Race",
+    Bias_Type == "ETHNICITY_BIAS" ~ "Ethnicity",
+    Bias_Type == "SEXUAL_ORIENTATION_BIAS" ~ "Sexual Orientation",
+    Bias_Type == "GENDER_BIAS" ~ "Gender",
+    Bias_Type == "LANGUAGE_BIAS" ~ "Language",
+    Bias_Type == "RELIGION_BIAS" ~ "Religion",
+    Bias_Type == "MENTAL_OR_PHYSICAL_DISABILITY" ~ "Disability",
+    TRUE ~ Bias_Type
+  ))
+
+# Create a bar chart
+ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(x = "Bias Type", y = "Count of Occurrences") +
+  scale_fill_brewer(palette = "Set3") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+@fig-occurrence-by-bias presents the total occurrences of hate crimes over the past six years, categorized by the type of bias involved.
+The data reveals that hate crimes motivated by ethnicity bias are the most frequently reported, underscoring significant concerns for ethnically targeted offenses within the community.
+In contrast, hate crimes related to disability bias are notably underreported.
+
+This discrepancy may not solely reflect the actual prevalence of such crimes; rather, it also highlights potential biases in measurement and reporting.
+While hate crimes based on ethnicity can often manifest as attacks on community spaces, disabled individuals frequently experience isolation, lacking the community structures that might support them in reporting incidents.
+This isolation can lead to underreporting of hate crimes, as many victims may not feel a sense of community or empowerment to come forward.
+Additionally, the unique barriers faced by individuals with disabilities, including fear of stigma and difficulties accessing reporting mechanisms, compound this issue.
+
+Additionally, the low reporting of hate crimes motivated by language bias also highlights potential biases in measurement and reporting.
+Many people may associate language with ethnicity, leading them to prioritize ethnicity over language in their perceptions of hate crimes.
+This tendency can obscure the realities faced by victims of language bias, further complicating the understanding of hate crime dynamics.
+
 ```{r}
 #| label: fig-religion-bias
-#| fig-cap: Occurrences of Hate Crimes by Religion Bias
+#| fig-cap: "Occurrences of Hate Crimes by Religion Bias"
 #| echo: false
+
 # Filter out "None" from the RELIGION_BIAS column and count occurrences
@@
 
+@fig-religion-bias illustrates the occurrences of hate crimes categorized by different religion bias types.
+The data reveals that, while a significant portion of hate crimes is directed toward individuals based on religion, Jewish individuals are disproportionately affected, facing the highest number of reported incidents.
+This trend underscores a critical concern regarding the targeting of Jewish communities, which highlights the ongoing issues of anti-Semitism within society.
+
 ```{r}
 #| label: fig-Jewish-bias
-#| fig-cap: Occurrences of Hate Crimes by Year and Race Bias
+#| fig-cap: "Trends in Hate Crimes by Religious Bias: Jewish vs. Other Religions"
 #| echo: false
+
 # Filter out "None" and create a new column for Race Bias categorization
@@
 ```
+@fig-Jewish-bias displays the percentage of hate crimes targeting Jewish individuals compared to other religious groups since 2018.
+The data reveals fluctuations in the proportion of hate crimes directed toward Jewish people.
+Specifically, the percentage has varied, with notable increases in certain years—rising from 70% in 2018 to 83% in 2022, followed by a slight decline to 78% in 2023.
+While this does not indicate a steady increase, it highlights consistent high levels of hate crimes against Jewish individuals, with occasional variations across the years.
+This trend suggests a sustained and significant targeting of the Jewish community.
 
 ```{r}
+#| label: fig-jewish-offence-location
+#| fig-cap: "Occurrences of Hate Crimes by Primary Offence for Jewish Bias"
 #| echo: false
-# Filter out "None" from the RACE_BIAS column and count occurrences
-race_bias_counts <- cleaned_data %>%
-  filter(RACE_BIAS != "None") %>%
-  count(RACE_BIAS)
 
-# Create the bar chart
-ggplot(race_bias_counts, aes(x = reorder(RACE_BIAS, n), y = n)) +
+# Filter out offenses for "Jewish" from the RELIGION_BIAS column and count occurrences
+jewish_offence_counts <- cleaned_data %>%
+  filter(RELIGION_BIAS == "Jewish") %>%
+  drop_na(LOCATION_TYPE) %>%
+  count(LOCATION_TYPE)
+
+# Create a named vector with the old and new labels
+location_labels <- c(
+  "Apartment Building (Condo, Retirement Buidling, etc.)" = "Apartment Building", 
+  "Business/Retail" = "Business/Retail", 
+  "Educational Institution (Universities, Colleges, Schools, etc.)" = "Educational Institution", 
+  "Government Building" = "Government Facility", 
+  "Government Building (Courthouse, Museums, Parliament Building, etc.)" = "Government Building", 
+  "House (Townhouse, Retirement Home, Garage, Vehicle, Cottage)" = "House", 
+  "Medical Facility (Hospitals, Long-term Care, etc.)" = "Medical Facility", 
+  "Non-Commercial/Non for Profit" = "Non-Commercial", 
+  "Open Area, Park or Parking Lot" = "Open Area", 
+  "Other Commercial / Corporate Places" = "Corporate Places", 
+  "Public Transportation" = "Public Transportation", 
+  "Religious Place of Worship/Cultural Centre" = "Religious Place of Worship/Cultural Centre", 
+  "Streets/Roadways/Highway" = "Streets/Roadways/Highway"
+
+  # Add more mappings here as needed
+)
+
+# Create the bar chart with renamed x-axis labels
+ggplot(jewish_offence_counts, aes(x = reorder(LOCATION_TYPE, n), y = n)) +
   geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
   theme_minimal() +
-  labs(x = "Race Bias Categories", y = "Number of Occurrences", 
-       title = "Occurrences of Hate Crimes by Race Bias") +
+  labs(x = "Location Type", y = "Number of Occurrences") +  # Rename x-axis label
+  scale_x_discrete(labels = location_labels) +  # Rename the x-axis categories
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 ```
 
-```{r}
-#| echo: false
-# Filter out "None" and create a new column for Race Bias categorization
-race_counts <- cleaned_data %>%
-  filter(RACE_BIAS != "None") %>%
-  mutate(RACE_CATEGORY = ifelse(RACE_BIAS == "Black", "Black", "Other"))
+@fig-jewish-offence-location illustrates the distribution of hate crimes against Jewish individuals across various locations.
+The data reveals that a significant portion of these offenses occur in educational institutions, particularly schools, indicating that Jewish students and staff may be especially vulnerable in these settings.
+This highlights a pressing concern regarding the safety and well-being of Jewish individuals within the educational environment.
 
-# Count occurrences by OCCURRENCE_YEAR and RACE_CATEGORY
-year_race_counts <- race_counts %>%
-  count(OCCURRENCE_YEAR, RACE_CATEGORY)
+In contrast, hate crimes reported at cultural centers are considerably less frequent.
+This suggests that these venues, which often function as community hubs for Jewish populations, may offer a safer atmosphere or are less likely to be targeted for such offenses.
 
-# Calculate percentage for each category within each year
-year_race_counts <- year_race_counts %>%
-  group_by(OCCURRENCE_YEAR) %>%
-  mutate(percentage = n / sum(n) * 100)
-
-# Create the stacked bar chart with percentage labels
-ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RACE_CATEGORY)) +
-  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
-  scale_fill_manual(values = c("Black" = "steelblue", "Other" = "orange")) +
-  theme_minimal() +
-  labs(x = "Year of Occurrence", y = "Percentage of Occurrences", 
-       title = "Percentage of Hate Crimes by Year and Race Bias") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_text(aes(label = scales::percent(percentage / 100)), 
-            position = position_stack(vjust = 0.5), 
-            color = "white", size = 4)  # Adjust size and color as needed
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
+The stark difference between the frequency of hate crimes in schools compared to cultural centers prompts critical questions regarding the factors that contribute to these disparities.
+Understanding the dynamics of these environments can guide targeted initiatives aimed at enhancing safety and raising awareness in high-risk areas, particularly schools, where Jewish individuals face heightened risks of victimization.
 

### 2024-09-27 scripts/00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates data for Hate Crimes analysis
 # Author: Sakura Hu
-# Date: 20 September 2024
+# Date: 27 September 2024
 # Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -21,15 +21,44 @@ n <- 100
 # Simulating OCCURRENCE_YEAR (e.g., from 2018 to 2023)
 occurrence_year <- sample(2018:2023, n, replace = TRUE)
 
+# Define proportions for each bias type
+gender_bias_levels <- unique(simulated_data$GENDER_BIAS)
+gender_bias_probs <- c(0.05, 0.7, 0.05, 0.1, 0.05, 0.05) 
+
+race_bias_levels <- unique(simulated_data$RACE_BIAS)
+race_bias_probs <- c(0.01, 0.01, 0.01, 0.01, 0.01, 0.3, 0.4, 0.01, 0.01, 0.01, 0.04, 0.03, 0.04, 0.04, 0.04, 0.02, 0.01)
+
+religion_bias_levels <- unique(simulated_data$RELIGION_BIAS)
+religion_bias_probs <- c(0.02, 0.02, 0.4, 0.02, 0.02, 0.02, 0.02, 0.4, 0.02, 0.02, 0.04)
+
+ethnicity_bias_levels <- unique(simulated_data$ETHNICITY_BIAS)
+ethnicity_bias_probs <- c(0.01, 0.01, 0.05, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.05, 0.01, 0.04, 0.6, 
+                          0.05, 0.05, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01)  
+
+sexual_orientation_bias_levels <- unique(simulated_data$SEXUAL_ORIENTATION_BIAS)
+sexual_orientation_bias_probs <- c(0.1, 0.1, 0.1, 0.6, 0.1)
+
+language_bias_levels <- unique(simulated_data$LANGUAGE_BIAS)
+language_bias_probs <- c(0.01, 0.01, 0.97, 0.01)
+
+mental_physical_disability_levels <- unique(simulated_data$MENTAL_OR_PHYSICAL_DISABILITY)
+mental_physical_disability_probs <- c(0.3, 0.7) 
+
+# Simulating the biases based on defined proportions
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from Open Data Toronto
 # Author: Sakura Hu
-# Date: 20 September 2024
+# Date: 27 September 2024
 # Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -34,14 +34,10 @@ data
 write_csv(data, "data/raw_data/raw_data.csv") 
 
 #### Save data ####
-# [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
 write_csv(data, "data/raw_data.csv") 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
 
 #### Save data ####
-# [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
 write_csv(raw_data, "inputs/data/raw_data.csv") 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw hate crime data recorded by observers
 # Author: Sakura Hu
-# Date: 23 September 2024
+# Date: 27 September 2024
 # Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: R packages tidyverse, dplyr must be installed
@@ -16,12 +16,47 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 # Replace "NO" with "None" in the specified columns
 cleaned_data <- raw_data %>%
   mutate(
-    AGE_BIAS = ifelse(AGE_BIAS == "NO", "None", AGE_BIAS),
     MENTAL_OR_PHYSICAL_DISABILITY = ifelse(MENTAL_OR_PHYSICAL_DISABILITY == "NO", "None", MENTAL_OR_PHYSICAL_DISABILITY)
   ) %>%
   # Remove specified columns
   select(-NEIGHBOURHOOD_140, -HOOD_140, -NEIGHBOURHOOD_158, -HOOD_158, -DIVISION, -REPORTED_TIME, 
-         -OCCURRENCE_TIME, -REPORTED_YEAR, -REPORTED_DATE, -AGE_BIAS, -MULTIPLE_BIAS, -LOCATION_TYPE)
+         -OCCURRENCE_TIME, -REPORTED_YEAR, -REPORTED_DATE, -MULTIPLE_BIAS, -'_id', -AGE_BIAS, 
+         -PRIMARY_OFFENCE, -ARREST_MADE, -OCCURRENCE_DATE)
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+
+#### Test data ####
+# 1. Check for missing values (NAs) in key columns
+na_checks <- sapply(cleaned_data, function(x) sum(is.na(x)))
+print("Number of NAs in each column:")
+print(na_checks)
+
+# 2. Test for uniqueness of EVENT_UNIQUE_ID (Assuming this column exists in your dataset)
+# Test for uniqueness of EVENT_UNIQUE_ID
+if ("EVENT_UNIQUE_ID" %in% colnames(cleaned_data)) {
+  # Identify duplicated EVENT_UNIQUE_IDs
+  duplicates <- cleaned_data$EVENT_UNIQUE_ID[duplicated(cleaned_data$EVENT_UNIQUE_ID) | 
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Sanity check of the data
 # Author: Sakura Hu
-# Date: 20 September 2024
+# Date: 27 September 2024
 # Contact: sakura.hu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have simulated data
@@ -15,12 +15,15 @@ library(tidyverse)
 #### Test data ####
 simulated_data <- read_csv("data/raw_data/simulated.csv")
 
-# Test for negative numbers
-simulated_data$Count |> min() <= 0
+# 1. Check for missing values (NAs) in key columns
+na_checks <- sapply(simulated_data, function(x) sum(is.na(x)))
+print("Number of NAs in each column:")
+print(na_checks)
 
-# Test for NAs
-all(is.na(simulated_data$Count))
+# 2. Test that all OCCURRENCE_YEAR values are between 2018 and 2023
+valid_years <- all(simulated_data$OCCURRENCE_YEAR %in% 2018:2023)
+print(paste("All OCCURRENCE_YEAR values are valid (2018 to 2023):", valid_years))
 
-# Test for unqiue id
-
-# Test for valid
+# 3. Check for duplicate rows
+duplicate_rows <- any(duplicated(simulated_data))
+print(paste("There are duplicate rows in the dataset:", duplicate_rows))

### 2024-09-27 paper/paper.qmd
final draft

@@
 title: "Analyzing Patterns of Hate Crimes in Toronto"
-subtitle: "majority of hate crimes are directed toward Jewish "
+subtitle: "The Majority of Hate Crimes Target Jewish Individuals"
 author: 
   - Sakura Hu
-thanks: "https://open.toronto.ca/dataset/hate-crimes-open-data/"
-date: "2024-09-27"
+thanks: "Code and data are available at: https://github.com/xycww/Hate_Crime_In_Toronto"
+date: "27 September 2024"
 date-format: long
@@
 
-Hate crimes are criminal offenses directed at individuals or property, driven by bias, prejudice, or hatred based on factors such as race, ethnicity, language, religion, gender, sexual orientation, disability, or other similar characteristics.
+Hate crimes are criminal offenses directed at individuals or property, driven by bias, prejudice, or hatred based on factors such as race, ethnicity, language, religion, gender, sexual orientation, disability, or other similar characteristics (Toronto Police Service, 2024).
 These crimes are a serious societal issue, as they target specific groups and reinforce divisions.
-In recent years, cities like Toronto have witnessed a rise in reported hate crimes, prompting a need to analyze the underlying causes and trends.
+In recent years, cities like Toronto have witnessed a rise in reported hate crimes, prompting a need to analyze the underlying causes and trends (The Canadian Press, 2024).
 Understanding the motivations behind hate crimes and identifying the most affected communities is critical for developing targeted prevention strategies and fostering a more inclusive society.
 
-This paper utilizes data on hate crimes reported in Toronto to explore the causes behind these offenses.
-Specifically, it investigates the most frequent motivations for hate crimes, including race, ethnicity, religion, sexual orientation, and gender identity.
-By analyzing which demographic groups are most often victimized, this paper aims to shed light on patterns of bias in hate crimes and how these trends have evolved over time.
-The analysis reveals that Black, Jewish, and female individuals are disproportionately targeted, and the occurrence of hate crimes has been increasing each year.
+This paper employs a dataset of reported hate crimes in Toronto to investigate the patterns and motivations behind these offenses.
+By analyzing demographic groups that are most frequently victimized, this research aims to uncover trends of bias and discrimination that persist in the community.
+The findings indicate that Jewish individuals are disproportionately targeted, with a concerning increase in hate crime occurrences each year.
+This pattern not only reflects existing societal prejudices but also raises alarms about the broader implications for community safety and cohesion.
 
-The importance of this study lies in its ability to inform policymakers, law enforcement, and the public about the prevalence and nature of hate crimes in Toronto.
-Addressing the root causes of these offenses is essential for reducing their occurrence and protecting vulnerable communities.
+The significance of this study lies in its potential to inform policymakers, law enforcement, and the general public about the prevalence and nature of hate crimes in Toronto.
+By addressing the root causes of these offenses, stakeholders can implement effective strategies to reduce their occurrence and protect vulnerable communities.
 
-The remainder of this paper is structured as follows: Section 2 discusses the data sources and variables, as well as the process of cleaning and preparing the data for analysis.
-Section 3 presents the results, including visualizations of key trends, while Section 4 provides a discussion of the findings.
-Finally, Section 5 includes appendices with additional details and supporting material.
+The remainder of this paper is structured as follows: @sec-data-overview discusses the data sources and variables used in the analysis; @sec-data-analysis presents the results, including visualizations of key trends and a discussion of the findings.
+@sec-data-cleaning in the appendix outlines the data cleaning process; @sec-reference include a list of reference.
 
 # Data {#sec-data}
-## Data Overview
+## Data Overview {#sec-data-overview}
 
-The dataset for this analysis is sourced from Open Data Toronto, specifically from the Toronto Police Service’s verified Hate Crime occurrences.
-It records all confirmed hate crimes that have been investigated by the Hate Crime Unit since 2018, categorized by their reported date.
-Notably, the dataset only includes verified hate crimes and excludes incidents that were unfounded or classified merely as hate incidents.
+The dataset used for this analysis is sourced from Open Data Toronto (Gelfand, 2022), specifically from the Toronto Police Service's verified hate crime occurrences.
+It records all confirmed hate crimes that have been investigated by the Hate Crime Unit since 2018, categorized by various types of bias.
+The dataset contains 1,350 observations and 25 variables.
+It is important to note that only verified hate crimes are included, while unfounded incidents or those classified merely as hate incidents are excluded.
 
 The dataset serves an important societal purpose, providing communities with detailed information to enhance public safety and awareness about hate crimes in Toronto.
-The data categories, referred to as bias categories, capture various dimensions of prejudice that are present in hate crimes, such as age, mental or physical disability, race, ethnicity, language, religion, sexual orientation, and gender.
+The data categories, referred to as bias categories, capture various dimensions of prejudice that are present in hate crimes.
 These categories allow researchers to analyze the prevalence of bias in different forms, offering a structured view of hate crime patterns over time.
 
-This dataset is invaluable for understanding hate crimes within the city, but it also comes with ethical considerations.
-Since each data point represents a real-world event, we must recognize the harm experienced by individuals and communities due to these crimes.
-Additionally, it’s important to acknowledge that the dataset focuses on confirmed hate crimes, which means it only represents cases where sufficient evidence was gathered.
-While this strengthens the reliability of the data, it also limits the scope of the analysis to known and proven crimes.
-Furthermore, the geographic locations of the incidents are aggregated to the neighborhood level to protect the privacy of the victims and communities involved.
+This dataset is essential for understanding hate crimes within the city, but it also comes with ethical considerations.
+Each data point represents a real-world event, requiring acknowledgment of the harm experienced by individuals and communities affected by these crimes.
+Additionally, the dataset only includes confirmed hate crimes, representing cases where sufficient evidence was gathered (Toronto Police Service, 2024).
+While this enhances the reliability of the data, it also limits the analysis to known and proven crimes.
+Furthermore, the geographic locations of the incidents are aggregated to the neighborhood level to protect the privacy of the victims and communities involved (Toronto Police Service, 2024).
 
-The data is composed of several key variables relevant to this analysis:
+The dataset was prepared, cleaned, and analyzed using R (R Core Team, 2022) with the following libraries: opendatatoronto (Gelfand, 2022) for accessing the data, tidyverse (Wickham et al., 2019), dplyr (Wickham et al., 2023) for data manipulation, and ggplot2 (Wickham, 2016) for visualizations.
+Additionally, knitr (Xie, 2023a) was used for report generation, and styler (Müller et al., 2024) ensured the R code was properly styled.
+
+The cleaned data that is used for analysis is composed of 10 variables:
 
@@
 - **MENTAL_OR_PHYSICAL_DISABILITY**: Whether the offense was based on the suspect’s perception of the victim's mental or physical disability.
-- **RACE_BIAS**: The race of the victim as perceived by the suspect, which was a basis for the hate crime.
-- **ETHNICITY_BIAS**: The ethnicity of the victim as perceived by the suspect, contributing to the hate crime.
-- **LANGUAGE_BIAS**: The language spoken by the victim as perceived by the suspect, used as a bias basis for the crime.
-- **RELIGION_BIAS**: The victim’s religion as perceived by the suspect, which motivated the crime.
-- **SEXUAL_ORIENTATION_BIAS**: The sexual orientation of the victim as perceived by the suspect, serving as a basis for the hate crime.
+- **RACE_BIAS**: The race of the victim as perceived by the suspect.
+- **ETHNICITY_BIAS**: The ethnicity of the victim as perceived by the suspect.
+- **LANGUAGE_BIAS**: The Language spoken by the victim as perceived by the suspect.
+- **RELIGION_BIAS**: The religion of the victim as perceived by the suspect.
+- **SEXUAL_ORIENTATION_BIAS**: The sexual orientation of the victim as perceived by the suspect.
 - **GENDER_BIAS**: The gender of the victim as perceived by the suspect.
-- **LOCATION_TYPE**: Location type of the offence.
+- **LOCATION_TYPE**: The type of location where the crime occurred (e.g., public space, residence).
 
-Alternative datasets were considered but were not used in this analysis.
-This particular dataset was chosen because it is published by the Toronto Police Service and provides verified, official data on hate crimes, making it more reliable than other possible datasets from unofficial sources.
-Including additional datasets could introduce inconsistency or unreliability, so this analysis focuses solely on this verified source.
+Alternative datasets were considered but ultimately not used.
+This dataset, published by the Toronto Police Service, provides verified, official data on hate crimes, making it more reliable than unofficial sources.
+The official nature of the data strengthens the credibility of the analysis, allowing for a more accurate reflection of hate crime occurrences and bias trends in Toronto.
 
-## Data Cleaning
-The dataset underwent several cleaning steps to ensure its relevance and consistency for analysis.
-Several variables were removed as they were either irrelevant to the current analysis or did not add value.
-These included neighborhood-level geographic identifiers (e.g., NEIGHBOURHOOD_140, NEIGHBOURHOOD_158), reporting times and dates (REPORTED_TIME, REPORTED_DATE), and any variables related to the classification or reporting of multiple biases.
-Specifically, AGE_BIAS was removed because it contained only the constant value “None,” suggesting that hate crimes in this dataset are rarely motivated solely by the victim’s age.
-
-Additionally, the MENTAL_OR_PHYSICAL_DISABILITY variable, which initially used "No" to denote a lack of bias, was modified to use "None" instead to maintain consistency with the other bias columns.
-No new variables were constructed, but this cleaning process ensured that the dataset was aligned and ready for analysis focused on the key bias categories.
-
-To handle the NAs in the location type column, rows with NAs location type is removed when doing relative analysis, it is not removed when analysis is irrelavant to location type.
-
-## Data Analysis
+## Data Analysis {#sec-data-analysis}
 
@@
   ) %>%
-  mutate(total_by_year = rowSums(across(starts_with("race_bias_count"):starts_with("disability_bias_count"))))
+  mutate(total_by_year = rowSums(across(
+    starts_with("race_bias_count"):starts_with("disability_bias_count")
+  )))
 
@@
   kable(
-    col.names = c("Year", "Race", "Ethnicity", "Religion", "Gender", "Sexual Orientation", "Disability", "Total"),
+    col.names = c(
+      "Year",
+      "Race",
+      "Ethnicity",
+      "Religion",
+      "Gender",
+      "Sexual Orientation",
+      "Disability",
+      "Total"
+    ),
     format = "pipe" # Use pipe format for compatibility
   )
-
 ```
@@
 Notably, incidents motivated by religious bias consistently represent the highest proportion among the various bias types, underscoring a significant area of concern for community safety and policy interventions.
-This data highlights not only the prevalence of hate crimes but also the specific biases that warrant focused attention and resources for prevention and support initiatives.
 
@@
 ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
+  geom_bar(stat = "identity",
+           fill = "steelblue",
+           alpha = 0.8) +
   theme_minimal() +
@@
 The data indicates a clear upward trend in the number of reported hate crimes over this period.
-Starting from approximately 154 occurrences in 2018, there is a steady increase each year, culminating in a significant rise to over 392 occurrences in 2023.
+Starting from approximately 154 occurrences in 2018, there is a steady increase each year, culminating in a significant rise to over 394 occurrences in 2023.
 
-This trend may suggest a growing prevalence of hate crimes within the community, possibly influenced by various social, political, and economic factors.
+This trend may suggest a growing prevalence of hate crimes within the Toronto, possibly influenced by various social, political, and economic factors.
 Additionally, it could reflect an increase in public awareness and reporting of such incidents, as societal attitudes toward hate crimes continue to evolve.
@@
 bias_counts <- cleaned_data %>%
-  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
-  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
+  select(
+    RACE_BIAS,
+    ETHNICITY_BIAS,
+    SEXUAL_ORIENTATION_BIAS,
+    GENDER_BIAS,
+    LANGUAGE_BIAS,
+    RELIGION_BIAS,
+    MENTAL_OR_PHYSICAL_DISABILITY
+  ) %>%
+  pivot_longer(cols = everything(),
+               names_to = "Bias_Type",
+               values_to = "Bias_Value") %>%
   mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
@@
 bias_counts <- bias_counts %>%
-  mutate(Bias_Type = case_when(
+  mutate(
+    Bias_Type = case_when(
       Bias_Type == "RACE_BIAS" ~ "Race",
@@
       TRUE ~ Bias_Type
-  ))
+    )
+  )
 
 # Create a bar chart
-ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
+ggplot(bias_counts, aes(
+  x = reorder(Bias_Type, Count),
+  y = Count,
+  fill = Bias_Type
+)) +
   geom_bar(stat = "identity") +
@@
 The data reveals that hate crimes motivated by ethnicity bias are the most frequently reported, underscoring significant concerns for ethnically targeted offenses within the community.
-In contrast, hate crimes related to disability bias are notably underreported.
+In contrast, hate crimes related to disability bias are notably under-reported.
 
 This discrepancy may not solely reflect the actual prevalence of such crimes; rather, it also highlights potential biases in measurement and reporting.
-While hate crimes based on ethnicity can often manifest as attacks on community spaces, disabled individuals frequently experience isolation, lacking the community structures that might support them in reporting incidents.
-This isolation can lead to underreporting of hate crimes, as many victims may not feel a sense of community or empowerment to come forward.
-Additionally, the unique barriers faced by individuals with disabilities, including fear of stigma and difficulties accessing reporting mechanisms, compound this issue.
+While hate crimes based on ethnicity can often manifest as attacks on community spaces, disabled individuals frequently experience isolation, lacking the community structures that might support them in reporting incidents (Healy, 2015).
+This isolation can lead to under-reporting of hate crimes, as many victims may not feel a sense of community or empowerment to come forward (Healy, 2015).
+Additionally, the unique barriers faced by individuals with disabilities, including fear of stigma and difficulties accessing reporting mechanisms, compound this issue (Healy, 2015).
 
@@
 ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
+  geom_bar(stat = "identity",
+           fill = "steelblue",
+           alpha = 0.8) +
   theme_minimal() +
@@
 # Create the stacked bar chart
-ggplot(year_religion_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RELIGION_CATEGORY)) +
-  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
+ggplot(year_religion_counts,
+       aes(
+         x = as.factor(OCCURRENCE_YEAR),
+         y = percentage,
+         fill = RELIGION_CATEGORY
+       )) +
+  geom_bar(stat = "identity",
+           position = "stack",
+           alpha = 0.8) +
   scale_fill_manual(values = c("Jewish" = "steelblue", "Other" = "orange")) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_text(aes(label = scales::percent(percentage / 100)), 
+  geom_text(
+    aes(label = scales::percent(percentage / 100)),
     position = position_stack(vjust = 0.5),
-            color = "white", size = 4)  # Adjust size and color as needed
+    color = "white",
+    size = 3
+  ) 
 ```
@@
 #| label: fig-jewish-offence-location
-#| fig-cap: "Occurrences of Hate Crimes by Primary Offence for Jewish Bias"
+#| fig-cap: "Distribution of Hate Crimes Against Jewish Individuals by Location Type"
 #| echo: false
@@
 ggplot(jewish_offence_counts, aes(x = reorder(LOCATION_TYPE, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
+  geom_bar(stat = "identity",
+           fill = "steelblue",
+           alpha = 0.8) +
   theme_minimal() +
@@
 
-# Results
+# Appendix {-}
 
-Our results are summarized in @tbl-modelresults.
+# Additional data details
 
-# Discussion
+## Data Cleaning {#sec-data-cleaning}
+The dataset underwent a series of cleaning steps to ensure its relevance and consistency for analysis.
+Several variables were removed due to their irrelevance to the current study.
+Notably, the AGE_BIAS variable was eliminated because it contained only the constant value “None,” suggesting that hate crimes are rarely motivated solely by the victim’s age.
 
-## First discussion point {#sec-first-point}
+Additionally, the MENTAL_OR_PHYSICAL_DISABILITY variable was updated to replace "No" with "None" to maintain consistency with the other bias categories.
+While no new variables were created, these cleaning processes ensured that the dataset was aligned and prepared for an analysis focused on the primary bias categories.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+To handle missing values in the location type column, rows with NAs in this variable were removed for analyses specifically related to location type.
+However, these rows were retained for analyses that did not pertain to location type, allowing for a more comprehensive examination of the data.
 
@@
 
-\appendix
+# References{#sec-reference}
+R Core Team. (2023). R: A language and environment for statistical computing. R Foundation for Statistical Computing. https://www.R-project.org/
 
-# Appendix {-}
+Gelfand, D. (2022). opendatatoronto: Access Open Data Toronto Datasets (R package version). https://CRAN.R-project.org/package=opendatatoronto
 
-\newpage
+Wickham, H., Averick, M., Bryan, J., Chang, W., D'Agostino McGowan, L., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T.
+L., Miller, E., Bache, S.
+M., Müller, K., Ooms, J., Robinson, D., Seidel, D., Spinu, V., Vaughan, K., Wilke, C., Takahashi, K., & Dunnington, D.
+(2019).
+Welcome to the tidyverse (R package version).
+https://CRAN.R-project.org/package=tidyverse
 
-# References
+Wickham, H., François, R., Henry, L., & Müller, K. (2023). dplyr: A grammar of data manipulation (R package version 1.1.2). https://CRAN.R-project.org/package=dplyr
+
+Wickham, H. (2016). ggplot2: Elegant graphics for data analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org
+
+Xie, Y. (2022). knitr: A general-purpose package for dynamic report generation in R (R package version 1.40). https://yihui.org/knitr/
+
+Müller, K., Walthert, L., & Patil, I. (2022). styler: Non-invasive pretty printing of R code (R package version 1.40). https://CRAN.R-project.org/package=styler
+
+Healy, J. (2015). Why don’t we recognise Disability Hate Crime? International Network for Hate Studies. https://internationalhatestudies.com/dont-recognise-disability-hate-crime/
+
+The Canadian Press. (2024). Hate crimes in Toronto up nearly 55% over last year: Police. CBC News. https://www.cbc.ca/news/canada/toronto/toronto-hate-crime-stats-1.7245301

### 2024-09-27 scripts/00-simulate_data.R

@@ -23,38 +23,95 @@ occurrence_year <- sample(2018:2023, n, replace = TRUE)
 
 # Define proportions for each bias type
 gender_bias_levels <- unique(simulated_data$GENDER_BIAS)
-gender_bias_probs <- c(0.05, 0.7, 0.05, 0.1, 0.05, 0.05) 
+# 70% None, 10% of women, and the others share the rest of proportion
+gender_bias_probs <- c(0.05, 0.7, 0.05, 0.1, 0.05, 0.05)
 
 race_bias_levels <- unique(simulated_data$RACE_BIAS)
-race_bias_probs <- c(0.01, 0.01, 0.01, 0.01, 0.01, 0.3, 0.4, 0.01, 0.01, 0.01, 0.04, 0.03, 0.04, 0.04, 0.04, 0.02, 0.01)
+# 40% None, 30% of Black, and the others share the rest of proportion, (Asians tend to share more)
+race_bias_probs <- c(
+  0.01,
+  0.01,
+  0.01,
+  0.01,
+  0.01,
+  0.3,
+  0.4,
+  0.01,
+  0.01,
+  0.01,
+  0.04,
+  0.03,
+  0.04,
+  0.04,
+  0.04,
+  0.02,
+  0.01
+)
 
 religion_bias_levels <- unique(simulated_data$RELIGION_BIAS)
+# 40% None, 40% of Jewish, and the others share the rest of proportion
 religion_bias_probs <- c(0.02, 0.02, 0.4, 0.02, 0.02, 0.02, 0.02, 0.4, 0.02, 0.02, 0.04)
 
 ethnicity_bias_levels <- unique(simulated_data$ETHNICITY_BIAS)
-ethnicity_bias_probs <- c(0.01, 0.01, 0.05, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.05, 0.01, 0.04, 0.6, 
-                          0.05, 0.05, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.01)  
+# 60% None, and the others share the rest of proportion. Asian and Indian will share more.
+ethnicity_bias_probs <- c(
+  0.01,
+  0.01,
+  0.05,
+  0.01,
+  0.01,
+  0.01,
+  0.01,
+  0.01,
+  0.01,
+  0.05,
+  0.01,
+  0.04,
+  0.6,
+  0.05,
+  0.05,
+  0.01,
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -28,16 +28,16 @@ resources <- list_package_resources("hate-crimes-open-data")
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data <- filter(datastore_resources, row_number() == 1) %>% get_resource()
 data
 #### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
+write_csv(data, "data/raw_data/raw_data.csv")
 
 #### Save data ####
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(data, "data/raw_data.csv") 
+write_csv(data, "data/raw_data.csv")
 
 
 #### Save data ####
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(raw_data, "inputs/data/raw_data.csv") 
+write_csv(raw_data, "inputs/data/raw_data.csv")

### 2024-09-27 scripts/02-data_cleaning.R

@@ -16,19 +16,26 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 # Replace "NO" with "None" in the specified columns
 cleaned_data <- raw_data %>%
   mutate(
-    MENTAL_OR_PHYSICAL_DISABILITY = ifelse(MENTAL_OR_PHYSICAL_DISABILITY == "NO", "None", MENTAL_OR_PHYSICAL_DISABILITY)
+    MENTAL_OR_PHYSICAL_DISABILITY = ifelse(
+      MENTAL_OR_PHYSICAL_DISABILITY == "NO",
+      "None",
+      MENTAL_OR_PHYSICAL_DISABILITY
+    )
   ) %>%
   # Remove specified columns
-  select(-NEIGHBOURHOOD_140, -HOOD_140, -NEIGHBOURHOOD_158, -HOOD_158, -DIVISION, -REPORTED_TIME, 
-         -OCCURRENCE_TIME, -REPORTED_YEAR, -REPORTED_DATE, -MULTIPLE_BIAS, -'_id', -AGE_BIAS, 
-         -PRIMARY_OFFENCE, -ARREST_MADE, -OCCURRENCE_DATE)
+  select(
+    -NEIGHBOURHOOD_140,-HOOD_140,-NEIGHBOURHOOD_158,-HOOD_158,-DIVISION,-REPORTED_TIME,
+    -OCCURRENCE_TIME,-REPORTED_YEAR,-REPORTED_DATE,-MULTIPLE_BIAS,-'_id',-AGE_BIAS,
+    -PRIMARY_OFFENCE,-ARREST_MADE,-OCCURRENCE_DATE
+  )
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
 
 #### Test data ####
 # 1. Check for missing values (NAs) in key columns
-na_checks <- sapply(cleaned_data, function(x) sum(is.na(x)))
+na_checks <- sapply(cleaned_data, function(x)
+  sum(is.na(x)))
 print("Number of NAs in each column:")
 print(na_checks)
 
@@ -36,7 +43,7 @@ print(na_checks)
 # Test for uniqueness of EVENT_UNIQUE_ID
 if ("EVENT_UNIQUE_ID" %in% colnames(cleaned_data)) {
   # Identify duplicated EVENT_UNIQUE_IDs
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -16,13 +16,17 @@ library(tidyverse)
 simulated_data <- read_csv("data/raw_data/simulated.csv")
 
 # 1. Check for missing values (NAs) in key columns
-na_checks <- sapply(simulated_data, function(x) sum(is.na(x)))
+na_checks <- sapply(simulated_data, function(x)
+  sum(is.na(x)))
 print("Number of NAs in each column:")
 print(na_checks)
 
 # 2. Test that all OCCURRENCE_YEAR values are between 2018 and 2023
 valid_years <- all(simulated_data$OCCURRENCE_YEAR %in% 2018:2023)
-print(paste("All OCCURRENCE_YEAR values are valid (2018 to 2023):", valid_years))
+print(paste(
+  "All OCCURRENCE_YEAR values are valid (2018 to 2023):",
+  valid_years
+))
 
 # 3. Check for duplicate rows
 duplicate_rows <- any(duplicated(simulated_data))
